---
title: "House Prices Project"
author: "Arjuna Anilkumar, A20446963"
date: "11/8/2020"
output: pdf_document
geometry: margin= 1cm
---


# Introduction 

This project aims to predict the final price of houses using the Ames housing dataset.

# Data description

The Ames Housing dataset was compiled by Dean De Cock for use in data science education. It's an alternative to the Boston Housing dataset and is for data scientists looking for a modernized and expanded version of the often cited Boston Housing dataset.

The Ames housing data contains With 79 explanatory variables describing (almost) every aspect of residential homes in Ames, Iowa.

# Data Processing 



## Install packages

```{r}
#install.packages(c("Amelia","purrr","tidyr","ggplot2","rpart","plyr","corrplot","RColorBrewer","ggrepel","DescTools","caret","xgboost","randomForest"))
```

## Load data

```{r}
df <- read.table("../data/raw/train.csv", sep = ",",header = T)
head(df)
combined <- df

combined <- combined[,!names(combined) %in% "Id"]
str(combined)
```


## Missing values

```{r}
library(Amelia)
misscounts <- sapply(combined,function(x) sum(is.na(x)))
missmap(combined, main = "Missing values")
```

```{r}
sort(colSums(sapply(combined, is.na)), decreasing = T)
```

### pool variables

The PoolQC has the most missing values. Pool area does not have missing values but it is related to PoolQC as it does not make sense to have a pool quality data when there is zero pool area or no pool. Its description from the data description document is.

PoolQC: Pool quality
		
       Ex	Excellent
       Gd	Good
       TA	Average/Typical
       Fa	Fair
       NA	No Pool

Since a house with no pool has NA they are not really missing values.
we can check with other pool related variables to see if there are any actual missing values in our data.

```{r}
table(is.na(combined$PoolQC))
table(combined$PoolArea, combined$PoolQC, useNA = 'ifany')
```

Here we have some actual missing values. We have 13 houses with pool area data but we have only 10 PoolQC data available. 


```{r}
library(plyr)
library(ggplot2)
combined[combined$PoolArea==0,]$PoolQC <- "None"

# convert all NA's in PoolQC to none except for the 3 actual missing values.

combined[is.na(combined$PoolQC),c("OverallQual","PoolArea")]

# imputing the values of poolQC according to overall quality and pool area.

combined[is.na(combined$PoolQC),"PoolQC"] <- c("TA","Gd","TA")

# label encoding as the values are ordinal.

quality <- c("None" = 0, 'Po' = 1 , 'Fa' = 2, 'TA'=3 , 'Gd'=4, 'Ex'=5 )

combined$PoolQC <- as.integer(revalue(combined$PoolQC, quality))

table(combined$PoolQC)
str(combined$PoolQC)
barplot(table(combined$PoolQC), xlab = "PoolQC", ylab = "Count")
ggplot(combined, aes(x=PoolQC, y = SalePrice)) + geom_bar(stat = 'summary')
```

```{r}
table(combined$PoolArea)
barplot(table(combined$PoolArea), xlab = "PoolArea", ylab = "Count")
ggplot(combined, aes(x=PoolArea, y = SalePrice)) + geom_point()
```

### MiscFeature variable 

```{r}
table(combined$MiscFeature, useNA = "ifany")
```

In MiscFeature variable, there are 1406 missing values that have to be replaced by none.

```{r}
library(ggplot2)

# convert all NA's in MiscFeature to none.
combined[is.na(combined$MiscFeature),"MiscFeature"] <- "None"

# convert to factor
combined$MiscFeature <- as.factor(combined$MiscFeature)

ggplot(combined, aes(x=MiscFeature, y = SalePrice)) + geom_bar(stat = 'summary')
```

### Alley Predictor

```{r}
table(combined$Alley, useNA = "ifany")
```

```{r}
# convert all NA's in Alley to none.
combined[is.na(combined$Alley),"Alley"] <- "None"

# convert to factor
combined$Alley <- as.factor(combined$Alley)

ggplot(combined, aes(x=Alley, y = SalePrice)) + geom_bar(stat = 'summary')
```

### Fence predictor


```{r}
table(combined$Fence, useNA = "ifany")
```

```{r}
# convert all NA's in Fence to none.
combined[is.na(combined$Fence),"Fence"] <- "None"

# convert to factor
combined$Fence <- as.factor(combined$Fence)

ggplot(combined, aes(x=Fence, y = SalePrice)) + geom_bar(stat = 'summary')
```

### Fireplace variables

Fireplace quality

```{r}
table(combined$FireplaceQu, useNA = "ifany")
```

```{r}
# convert all NA's in FireplaceQu to none.

combined[is.na(combined$FireplaceQu),"FireplaceQu"] <- "None"

# Changing and converting to factor levels from character.


combined$FireplaceQu <- as.integer(revalue(combined$FireplaceQu, quality))

table(combined$FireplaceQu, useNA = "ifany")
str(combined$FireplaceQu)
anyNA(combined$FireplaceQu)

barplot(table(combined$FireplaceQu), xlab = "FireplaceQu", ylab = "Count")
ggplot(combined, aes(x=FireplaceQu, y = SalePrice)) + geom_bar(stat = 'summary')
```

### Lot variables

LotFrontage
LotShape
LotConfig
LotArea

```{r}
table(is.na(combined$LotFrontage))
```

Here we have 259 missing values which cannot be replaced by none as it is a numerical variable. So we predict using rpart.

http://r-statistics.co/Missing-Value-Treatment-With-R.html

```{r}
# predictors that lotfrontage variable might depend on.
predictors <- c("MSSubClass", "MSZoning", "LotFrontage", "LotArea", "Street", "Alley", "LotShape", "LandContour", "LotConfig", "LandSlope", "BldgType", "HouseStyle", "YrSold", "SaleType", "SaleCondition")
library(rpart)
mod <- rpart(LotFrontage~., data = combined[!is.na(combined$LotFrontage),predictors], method = "anova", na.action = na.omit)

pred <- predict(mod, combined[is.na(combined$LotFrontage),predictors])
pred <- round(pred)
combined$LotFrontage[is.na(combined$LotFrontage)] <- pred
anyNA(combined$LotFrontage)
ggplot(combined, aes(x=LotFrontage, y = SalePrice)) + geom_point()
```

```{r}
table(combined$LotShape, useNA = "ifany")

combined$LotShape <- as.integer(revalue(combined$LotShape,  c("IR3"=0 , "IR2"=1 , "IR1"=2 , "Reg"=3 )))
barplot(table(combined$LotShape), xlab = "LotShape", ylab = "Count")
ggplot(combined, aes(x=LotShape, y = SalePrice)) + geom_bar(stat = 'summary')
```

```{r}
table(combined$LotConfig, useNA = "ifany")
combined$LotConfig <- as.factor(combined$LotConfig)
barplot(table(combined$LotConfig), xlab = "LotConfig", ylab = "Count")
ggplot(combined, aes(x=LotConfig, y = SalePrice)) + geom_bar(stat = 'summary')
```

###  Garage variables 

GarageYrBlt
GarageType
GarageFinish,
GarageQual,
GarageCond,
GarageCars,
GarageArea

```{r}
garage <- c("GarageYrBlt","GarageType","GarageFinish","GarageQual","GarageCond","GarageCars","GarageArea")
sort(colSums(sapply(combined[,garage], is.na)), decreasing = T)

```

```{r}
combined$GarageYrBlt[is.na(combined$GarageYrBlt)] <- combined$YearBuilt[is.na(combined$GarageYrBlt)]

combined$GarageType[is.na(combined$GarageType)] <- "None"
combined$GarageFinish[is.na(combined$GarageFinish)] <- "None"
combined$GarageCond[is.na(combined$GarageCond)] <- "None"
combined$GarageQual[is.na(combined$GarageQual)] <- "None"
sort(colSums(sapply(combined[,garage], is.na)), decreasing = T)
```


```{r}
# convert into factor
combined$GarageType <- as.factor(combined$GarageType)
table(combined$GarageType)
barplot(table(combined$GarageType), xlab = "GarageType", ylab = "Count")
ggplot(combined, aes(x=GarageType, y = SalePrice)) + geom_bar(stat = 'summary')
```

```{r}
Finish <- c('None'=0, 'Unf'=1, 'RFn'=2, 'Fin'=3)
combined$GarageFinish<-as.integer(revalue(combined$GarageFinish,  Finish))
table(combined$GarageFinish, useNA = 'ifany')
barplot(table(combined$GarageFinish), xlab = "GarageFinish", ylab = "Count")
ggplot(combined, aes(x=GarageFinish, y = SalePrice)) + geom_bar(stat = 'summary')
```

```{r}
combined$GarageCond<-as.integer(revalue(combined$GarageCond,  quality))
table(combined$GarageCond, useNA = "ifany")
barplot(table(combined$GarageCond), xlab = "GarageCond", ylab = "Count")
ggplot(combined, aes(x=GarageCond, y = SalePrice)) + geom_bar(stat = 'summary')
```

```{r}
combined$GarageQual<-as.integer(revalue(combined$GarageQual,  quality))
table(combined$GarageQual, useNA = "ifany")
barplot(table(combined$GarageQual), xlab = "GarageQual", ylab = "Count")
ggplot(combined, aes(x=GarageQual, y = SalePrice)) + geom_bar(stat = 'summary')
```

```{r}

barplot(table(combined$GarageCars), xlab = "GarageCars", ylab = "Count")
ggplot(combined, aes(x=GarageCars, y = SalePrice)) + geom_bar(stat = 'summary')
```

```{r}
ggplot(combined, aes(x=GarageArea, y = SalePrice)) + geom_point()
```

### Basement variables

there are 11 basement variables

BsmtQual,
BsmtCond,
BsmtExposure,
BsmtFinType1,
BsmtFinType2,
BsmtFullBath,
BsmtHalfBath,
BsmtFinSF1,
BsmtFinSF2,
BsmtUnfSF,
TotalBsmtSF

```{r}
basement <- c("BsmtQual","BsmtCond","BsmtExposure","BsmtFinType1","BsmtFinType2","BsmtFullBath","BsmtHalfBath","BsmtFinSF1","BsmtFinSF2","BsmtUnfSF","TotalBsmtSF")
sort(colSums(sapply(combined[,basement], is.na)), decreasing = T)

```

```{r}
x <- which(!is.na(combined$BsmtFinType1) & (is.na(combined$BsmtCond)|is.na(combined$BsmtExposure)|is.na(combined$BsmtQual)|is.na(combined$BsmtFinType2)))

combined[x,basement]

# impute mode
combined[c(949),"BsmtExposure"] <- names(sort(-table(combined$BsmtExposure)))[1]
combined[c(333),"BsmtFinType2"] <- names(sort(-table(combined$BsmtFinType2)))[1]
combined[x,basement]
anyNA(combined[x,basement])
sort(colSums(sapply(combined[,basement], is.na)), decreasing = T)
```

```{r}
combined$BsmtQual[is.na(combined$BsmtQual)] <- "None"
combined$BsmtCond[is.na(combined$BsmtCond)] <- "None"
combined$BsmtExposure[is.na(combined$BsmtExposure)] <- "None"
combined$BsmtFinType1[is.na(combined$BsmtFinType1)] <- "None"
combined$BsmtFinType2[is.na(combined$BsmtFinType2)] <- "None"
combined$BsmtFullBath[is.na(combined$BsmtFullBath)] <- 0
combined$BsmtHalfBath[is.na(combined$BsmtHalfBath)] <- 0

sort(colSums(sapply(combined[,basement], is.na)), decreasing = T)
```

#### BsmtQual

```{r}
# label encoding

combined$BsmtQual<-as.integer(revalue(combined$BsmtQual,  quality))
table(combined$BsmtQual, useNA = "ifany")
barplot(table(combined$BsmtQual), xlab = "BsmtQual", ylab = "Count")
ggplot(combined, aes(x=BsmtQual, y = SalePrice)) + geom_bar(stat = 'summary')
```

#### BsmtCond

```{r}
combined$BsmtCond<-as.integer(revalue(combined$BsmtCond,  quality))
table(combined$BsmtCond, useNA = "ifany")
barplot(table(combined$BsmtCond), xlab = "BsmtCond", ylab = "Count")
ggplot(combined, aes(x=BsmtCond, y = SalePrice)) + geom_bar(stat = 'summary')
```

#### BsmtExposure

```{r}
exposure <- c('None'=0,'No'=1,'Mn'=2,'Av'=3,'Gd'=4)
combined$BsmtExposure<-as.integer(revalue(combined$BsmtExposure,  exposure))
table(combined$BsmtExposure, useNA = "ifany")
barplot(table(combined$BsmtExposure), xlab = "BsmtExposure", ylab = "Count")
ggplot(combined, aes(x=BsmtExposure, y = SalePrice)) + geom_bar(stat = 'summary')
```

#### BsmtFinType1

```{r}
rating <- c('None'=0,'Unf'=1,'LwQ'=2,'Rec'=3,'BLQ'=4,'ALQ'=5,'GLQ'=6)
combined$BsmtFinType1<-as.integer(revalue(combined$BsmtFinType1,  rating))
table(combined$BsmtFinType1, useNA = "ifany")
barplot(table(combined$BsmtFinType1), xlab = "BsmtFinType1", ylab = "Count")
ggplot(combined, aes(x=BsmtFinType1, y = SalePrice)) + geom_bar(stat = 'summary')
```

#### BsmtFinType2

```{r}
combined$BsmtFinType2<-as.integer(revalue(combined$BsmtFinType2,  rating))
table(combined$BsmtFinType2, useNA = "ifany")
barplot(table(combined$BsmtFinType2), xlab = "BsmtFinType2", ylab = "Count")
ggplot(combined, aes(x=BsmtFinType2, y = SalePrice)) + geom_bar(stat = 'summary')
```

#### BsmtFullBath

```{r}
barplot(table(combined$BsmtFullBath), xlab = "BsmtFullBath", ylab = "Count")
ggplot(combined, aes(x=BsmtFullBath, y = SalePrice)) + geom_bar(stat = 'summary')
```

#### BsmtHalfBath

```{r}
barplot(table(combined$BsmtHalfBath), xlab = "BsmtHalfBath", ylab = "Count")
ggplot(combined, aes(x=BsmtHalfBath, y = SalePrice)) + geom_bar(stat = 'summary')
```

#### BsmtFinSF1

```{r}
ggplot(combined, aes(x=BsmtFinSF1, y = SalePrice)) + geom_point()
```

#### BsmtFinSF2
```{r}
ggplot(combined, aes(x=BsmtFinSF2, y = SalePrice)) + geom_point()
```

#### BsmtUnfSF
```{r}
ggplot(combined, aes(x=BsmtUnfSF, y = SalePrice)) + geom_point()

```

#### TotalBsmtSF
```{r}
ggplot(combined, aes(x=TotalBsmtSF, y = SalePrice)) + geom_point()

```

### masonry variables
```{r}
table(combined$MasVnrType)
combined$MasVnrType[is.na(combined$MasVnrType)] <- "None"
combined$MasVnrArea[is.na(combined$MasVnrArea)] <- 0
combined$MasVnrType <- as.factor(combined$MasVnrType)
table(combined$MasVnrType)
barplot(table(combined$MasVnrType), xlab = "MasVnrType", ylab = "Count")
ggplot(combined, aes(x=MasVnrType, y = SalePrice)) + geom_bar(stat = 'summary')
```

### Electrical variable

```{r}
table(combined$Electrical)
combined$Electrical[is.na(combined$Electrical)] <- names(sort(-table(combined$Electrical)))[1]

combined$Electrical <- as.factor(combined$Electrical)
table(combined$Electrical)
barplot(table(combined$Electrical), xlab = "Electrical", ylab = "Count")
ggplot(combined, aes(x=Electrical, y = SalePrice)) + geom_bar(stat = 'summary')
```

### MSZoning

```{r}
table(combined$MSZoning)
combined$MSZoning <- as.factor(combined$MSZoning)
table(combined$MSZoning)
barplot(table(combined$MSZoning), xlab = "MSZoning", ylab = "Count")
ggplot(combined, aes(x=MSZoning, y = SalePrice)) + geom_bar(stat = 'summary')
```

### Street

```{r}
table(combined$Street)
combined$Street<-as.integer(revalue(combined$Street, c('Grvl'=0, 'Pave'=1)))
table(combined$Street)
barplot(table(combined$Street), xlab = "Street", ylab = "Count")
ggplot(combined, aes(x=Street, y = SalePrice)) + geom_bar(stat = 'summary')
```

### LandContour 

Lvl  Near Flat/Level 
Bnk  Banked - Quick and significant rise from street grade to building
HLS  Hillside - Significant slope from side to side
Low  Depression

```{r}
table(combined$LandContour)
combined$LandContour <- as.factor(combined$LandContour)
table(combined$LandContour)
barplot(table(combined$LandContour), xlab = "LandContour", ylab = "Count")
ggplot(combined, aes(x=LandContour, y = SalePrice)) + geom_bar(stat = 'summary')
```

### LandSlope

   Gtl  Gentle slope
   Mod  Moderate Slope  
   Sev  Severe Slope
```{r}
table(combined$LandSlope)
combined$LandSlope<-as.integer(revalue(combined$LandSlope, c('Sev'=0, 'Mod'=1, 'Gtl'=2)))
table(combined$LandSlope)
barplot(table(combined$LandSlope), xlab = "LandSlope", ylab = "Count")
ggplot(combined, aes(x=LandSlope, y = SalePrice)) + geom_bar(stat = 'summary')
```

### Utilities

```{r}
table(combined$Utilities)

# remove utilities as it does not give any valuable information.
combined <- combined[,!names(combined) %in% "Utilities"]
```

### Neighborhood
   
```{r}
table(combined$Neighborhood)
combined$Neighborhood <- as.factor(combined$Neighborhood)
table(combined$Neighborhood)
barplot(table(combined$Neighborhood), xlab = "Neighborhood", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=Neighborhood, y = SalePrice)) + geom_bar(stat = 'summary')

```

### Condition1 and Condition2

```{r}
table(combined$Condition1)
combined$Condition1 <- as.factor(combined$Condition1)
table(combined$Condition1)
barplot(table(combined$Condition1), xlab = "Condition1", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=Condition1, y = SalePrice)) + geom_bar(stat = 'summary')
```

```{r}
table(combined$Condition2)
combined$Condition2 <- as.factor(combined$Condition2)
table(combined$Condition2)
barplot(table(combined$Condition2), xlab = "Condition2", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=Condition2, y = SalePrice)) + geom_bar(stat = 'summary')
```

### BldgType

```{r}
table(combined$BldgType)
combined$BldgType <- as.factor(combined$BldgType)
table(combined$BldgType)
barplot(table(combined$BldgType), xlab = "BldgType", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=BldgType, y = SalePrice)) + geom_bar(stat = 'summary')
```

### HouseStyle

```{r}
table(combined$HouseStyle)
combined$HouseStyle <- as.factor(combined$HouseStyle)
table(combined$HouseStyle)
barplot(table(combined$HouseStyle), xlab = "HouseStyle", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=HouseStyle, y = SalePrice)) + geom_bar(stat = 'summary')
```

### RoofStyle

```{r}
table(combined$RoofStyle)
combined$RoofStyle <- as.factor(combined$RoofStyle)
table(combined$RoofStyle)
barplot(table(combined$RoofStyle), xlab = "RoofStyle", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=RoofStyle, y = SalePrice)) + geom_bar(stat = 'summary')
```

### RoofMatl

```{r}
table(combined$RoofMatl)
combined$RoofMatl <- as.factor(combined$RoofMatl)
table(combined$RoofMatl)
barplot(table(combined$RoofMatl), xlab = "RoofMatl", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=RoofMatl, y = SalePrice)) + geom_bar(stat = 'summary')
```

### Exterior variables
"Exterior1st"   "Exterior2nd"   "ExterQual"     "ExterCond"
```{r}
table(combined$Exterior1st)
combined$Exterior1st <- as.factor(combined$Exterior1st)
table(combined$Exterior1st)
barplot(table(combined$Exterior1st), xlab = "Exterior1st", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=Exterior1st, y = SalePrice)) + geom_bar(stat = 'summary')
```

```{r}
table(combined$Exterior2nd)
combined$Exterior2nd <- as.factor(combined$Exterior2nd)
table(combined$Exterior2nd)
barplot(table(combined$Exterior2nd), xlab = "Exterior2nd", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=Exterior2nd, y = SalePrice)) + geom_bar(stat = 'summary')
```

```{r}
table(combined$ExterQual)
combined$ExterQual<-as.integer(revalue(combined$ExterQual, quality))
table(combined$ExterQual)
barplot(table(combined$ExterQual), xlab = "ExterQual", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=ExterQual, y = SalePrice)) + geom_bar(stat = 'summary')
```

```{r}
table(combined$ExterCond)
combined$ExterCond<-as.integer(revalue(combined$ExterCond, quality))
table(combined$ExterCond)
barplot(table(combined$ExterCond), xlab = "ExterCond", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=ExterCond, y = SalePrice)) + geom_bar(stat = 'summary')
```

### Foundation

```{r}
table(combined$Foundation)
combined$Foundation <- as.factor(combined$Foundation)
table(combined$Foundation)
barplot(table(combined$Foundation), xlab = "Foundation", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=Foundation, y = SalePrice)) + geom_bar(stat = 'summary')
```

### heating variables
 "Heating"       "HeatingQC" 
 
```{r}
table(combined$Heating)
combined$Heating <- as.factor(combined$Heating)
table(combined$Heating)
barplot(table(combined$Heating), xlab = "Heating", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=Heating, y = SalePrice)) + geom_bar(stat = 'summary')
```

```{r}
table(combined$HeatingQC)
combined$HeatingQC<-as.integer(revalue(combined$HeatingQC, quality))
table(combined$HeatingQC)
barplot(table(combined$HeatingQC), xlab = "HeatingQC", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=HeatingQC, y = SalePrice)) + geom_bar(stat = 'summary')
```

### CentralAir

```{r}
table(combined$CentralAir)
combined$CentralAir<-as.integer(revalue(combined$CentralAir, c('N'=0, 'Y'=1)))
table(combined$CentralAir)
barplot(table(combined$CentralAir), xlab = "CentralAir", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=CentralAir, y = SalePrice)) + geom_bar(stat = 'summary')
```

### KitchenQual

```{r}
table(combined$KitchenQual)
combined$KitchenQual<-as.integer(revalue(combined$KitchenQual, quality))
table(combined$KitchenQual)
barplot(table(combined$KitchenQual), xlab = "KitchenQual", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=KitchenQual, y = SalePrice)) + geom_bar(stat = 'summary')
```

### Functional

```{r}
table(combined$Functional)
combined$Functional<-as.integer(revalue(combined$Functional, c('Sal'=0, 'Sev'=1, 'Maj2'=2, 'Maj1'=3, 'Mod'=4, 'Min2'=5, 'Min1'=6, 'Typ'=7)))
table(combined$Functional)
barplot(table(combined$Functional), xlab = "Functional", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=Functional, y = SalePrice)) + geom_bar(stat = 'summary')
```

### PavedDrive

```{r}
table(combined$PavedDrive)
combined$PavedDrive<-as.integer(revalue(combined$PavedDrive, c('N'=0, 'P'=1, 'Y'=2)))
table(combined$PavedDrive)
barplot(table(combined$PavedDrive), xlab = "PavedDrive", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=PavedDrive, y = SalePrice)) + geom_bar(stat = 'summary')
```

### Sales variables

 "SaleType"      "SaleCondition"
 
```{r}
table(combined$SaleType)
combined$SaleType <- as.factor(combined$SaleType)
table(combined$SaleType)
barplot(table(combined$SaleType), xlab = "SaleType", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=SaleType, y = SalePrice)) + geom_bar(stat = 'summary')
```

```{r}
table(combined$SaleCondition)
combined$SaleCondition <- as.factor(combined$SaleCondition)
table(combined$SaleCondition)
barplot(table(combined$SaleCondition), xlab = "SaleCondition", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=SaleCondition, y = SalePrice)) + geom_bar(stat = 'summary')
```

### Remaining character variables

```{r}
Charcol <- names(combined[,sapply(combined, is.character)])
Charcol
```

### MSSubClass

    20  1-STORY 1946 & NEWER ALL STYLES
    30  1-STORY 1945 & OLDER
    40  1-STORY W/FINISHED ATTIC ALL AGES
    45  1-1/2 STORY - UNFINISHED ALL AGES
    50  1-1/2 STORY FINISHED ALL AGES
    60  2-STORY 1946 & NEWER
    70  2-STORY 1945 & OLDER
    75  2-1/2 STORY ALL AGES
    80  SPLIT OR MULTI-LEVEL
    85  SPLIT FOYER
    90  DUPLEX - ALL STYLES AND AGES
   120  1-STORY PUD (Planned Unit Development) - 1946 & NEWER
   150  1-1/2 STORY PUD - ALL AGES
   160  2-STORY PUD - 1946 & NEWER
   180  PUD - MULTILEVEL - INCL SPLIT LEV/FOYER
   190  2 FAMILY CONVERSION - ALL STYLES AND AGES
   
```{r}
table(combined$MSSubClass)
combined$MSSubClass <- as.factor(combined$MSSubClass)

#revalue for better readability
combined$MSSubClass<-revalue(combined$MSSubClass, c('20'='1 story 1946+', '30'='1 story 1945-', '40'='1 story unf attic', '45'='1,5 story unf', '50'='1,5 story fin', '60'='2 story 1946+', '70'='2 story 1945-', '75'='2,5 story all ages', '80'='split/multi level', '85'='split foyer', '90'='duplex all style/age', '120'='1 story PUD 1946+', '150'='1,5 story PUD all', '160'='2 story PUD 1946+', '180'='PUD multilevel', '190'='2 family conversion'))
barplot(table(combined$MSSubClass), xlab = "MSSubClass", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=MSSubClass, y = SalePrice)) + geom_bar(stat = 'summary')
```

### MoSold
```{r}
table(combined$MoSold)
combined$MoSold <- as.factor(combined$MoSold)
table(combined$MoSold)
barplot(table(combined$MoSold), xlab = "MoSold", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=MoSold, y = SalePrice)) + geom_bar(stat = 'summary')
```

### YrSold
```{r}
table(combined$YrSold)
combined$YrSold <- as.factor(combined$YrSold)
table(combined$YrSold)
barplot(table(combined$YrSold), xlab = "YrSold", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=YrSold, y = SalePrice)) + geom_bar(stat = 'summary')
```

```{r}
sort(colSums(sapply(combined, is.na)), decreasing = T)
```

```{r}
misscounts <- sapply(combined,function(x) sum(is.na(x)))
missmap(combined, main = "Missing values")

anyNA(combined)
```

As we can see there are no missing values.

```{r}
num_vars <- names(Filter(is.numeric,combined)) #index vector numeric variables
factor_vars <- names(Filter(is.factor,combined))
cat('numeric variables: ', length(num_vars),' and categorical variables: ',length(factor_vars),'\n')
write.csv(combined, "../data/processed/clean_data.csv")
```


# EDA

## Summarize Datasets
```{r}
# primary dataset
str(combined)
```

## Numerical variables

### SalePrice variable
```{r}
library(ggplot2)
layout(mat = matrix(c(1,2),2,1, byrow=TRUE), height = c(1,8))
par(mar=c(0, 3.1, 1.1, 2.1))
boxplot(combined$SalePrice , horizontal=TRUE , xaxt="n", frame=F, main=sprintf('Histogram of SalePrice'))
par(mar=c(4, 3.1, 1.1, 2.1))
hist(combined$SalePrice,main='', xlab = "SalePrice", ylab = "count")
```

### SalePrice vs all other numerical variable scatterplots
```{r}
plot = function(variable){
  print(variable)
  ggplot(combined,aes(x = combined[,variable], y = SalePrice)) + geom_point() + theme_classic() + labs(x=variable)}
library(gridExtra)
p = list()
p <- NULL
val <- 0
d <- combined[,num_vars]
for(j in 1:6){
  for(i in 1:9){
    name = names(d[i+val])
    p[[i]] = plot(as.character(name))}
  val = i+val
  do.call(grid.arrange,p)
  p <- NULL}
```

### all numerical variables density plots

```{r}
library(purrr)
library(tidyr)
library(ggplot2)
combined %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) + 
      facet_wrap(~ key, scales = "free") + 
      geom_histogram(aes(y=..density..), fill = "grey") + 
      geom_density()
ggsave("../Plots/density_numerical_variables.jpg", plot = last_plot(), width = 10, height = 7)
```



## Categorical variables

### Histograms for Categorical Variables
```{r}
plot_factor = function(variable){
  ggplot(combined,aes(x = combined[,variable], y = SalePrice)) +geom_bar(stat = 'summary') + theme_classic() + labs(x=variable)}

library(gridExtra)
p = list()
p <- NULL
val <- 0
d <- combined[,factor_vars]
for(j in 1:5){
  for(i in 1:5){
    name = names(d[i+val])
    p[[i]] = plot_factor((name))}
  val = i+val
  do.call(grid.arrange,p)
  p <- NULL}
```

# Feature engineering

## Total bathrooms
```{r}
combined$TotBathrooms <- combined$FullBath + (combined$HalfBath*0.5) + combined$BsmtFullBath + (combined$BsmtHalfBath*0.5)
table(combined$TotBathrooms)
barplot(table(combined$TotBathrooms), xlab = "TotBathrooms", ylab = "Count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=TotBathrooms, y = SalePrice)) + geom_bar(stat = 'summary')
```

## House age and remodelled houses

After looking at the density plots and EDA, we found that there is no house age but ther e is remodeled year and year built variables so adding the age of the house to the data set and whether or not its been remodeled was important as Sale price does depend on how old the house is and whether or not its a remodeled house.

### house age plot
```{r}
combined$Remod <- ifelse(combined$YearBuilt==combined$YearRemodAdd, 0, 1) #0=No Remodeling, 1=Remodeling
combined$Age <- as.numeric(combined$YrSold)-combined$YearRemodAdd
ggplot(data=combined[!is.na(combined$SalePrice),], aes(x=Age, y=SalePrice))+
        geom_point(col='blue')+
        scale_y_continuous(breaks= seq(0, 800000, by=100000))
```

### histogram of remodeled house variable
```{r}
barplot(table(combined$Remod), xlab = "Remodeled", ylab = "count")
ggplot(combined[!is.na(combined$SalePrice),], aes(x=Remod, y = SalePrice)) + geom_bar(stat = 'summary')
```

# Correlations

## correlation matrix
```{r}
num_vars <- names(Filter(is.numeric,combined)) #index vector numeric variables
factor_vars <- names(Filter(is.factor,combined))

corr_matrix <- cor(combined[num_vars],method = "pearson")
correlation_SalePrice <- sort(corr_matrix[,'SalePrice'],decreasing = TRUE)
correlation_SalePrice
```

## correlation Plot
```{r}
library(corrplot)
corrplot(corr_matrix,type = "upper",tl.col = "black",tl.cex = 0.8)
```

## correlation Plot of variables vs SalePrice > 0.5
```{r}
corr_sorted <- as.matrix(sort(corr_matrix[,'SalePrice'], decreasing = TRUE))
 #select only high correlations
Corr_High <- names(which(apply(corr_sorted, 1, function(x) abs(x)>0.5)))
corr_matrix <- corr_matrix[Corr_High, Corr_High]

corrplot.mixed(corr_matrix, tl.col="black", tl.pos = "lt", tl.cex = 0.8,cl.cex = 0.8, number.cex=0.8)
```

## Removing highly correlated variables
```{r}
# removing variables with high correlations between predictor # variables
dropVars <- c('GarageYrBlt', 'GarageArea', 'TotalBsmtSF', 'TotalRmsAbvGrd')
combined <- combined[,!names(combined) %in% dropVars]
```

# Outlier analysis

## Detecting outliers

```{r}
# we look for outliers in the most correlated variable with SalePrice which is GrLivArea.
library(ggrepel)
ggplot(combined[!is.na(combined$SalePrice),], aes(x=GrLivArea  , y = SalePrice)) + geom_point()+geom_smooth(method = "lm", se=FALSE, color="black", aes(group=1)) + geom_text_repel(aes(label = ifelse(GrLivArea>4500, rownames(combined), '')))
```

```{r}
ggplot(combined[!is.na(combined$SalePrice),], aes(x=X1stFlrSF, y = SalePrice)) + geom_point()+geom_smooth(method = "lm", se=FALSE, color="black", aes(group=1)) + geom_text_repel(aes(label = ifelse(combined$X1stFlrSF>3000, rownames(combined), '')))
```

```{r}
mod <- lm(SalePrice ~ ., data = combined)
summary(mod)
cooksd <- cooks.distance(mod)
```

```{r}
# All outliers
influential <- as.numeric(names(cooksd)[(cooksd > 10*mean(cooksd, na.rm=T))])  # influential row numbers
influential <- na.omit(influential)
influential
```

## Deleting outliers
```{r}
combined <- combined[-influential,]
```

# Preparing data for modelling

Methods used in:
https://topepo.github.io/caret/pre-processing.html

```{r}
df <- df[,!names(df) %in% "Id"]
true_num_vars <- names(Filter(is.numeric,df))
num_vars <- names(Filter(is.numeric,combined)) #index vector numeric variables
factor_vars <- names(Filter(is.factor,combined))
cat('numeric variables: ', length(num_vars),' and categorical variables:',length(factor_vars),'\n')
```

## Standardizing numerical predictor variables
```{r}
library(caret)
library(DescTools)
nvarnames <- true_num_vars[!(true_num_vars %in% c('MSSubClass', 'MoSold', 'YrSold', 'SalePrice', 'OverallQual', 'OverallCond'))] 
num_df <- combined[, names(combined) %in% nvarnames]
factor_df <- combined[, !(names(combined) %in% nvarnames)]
factor_df <- factor_df[, names(factor_df) != 'SalePrice']

cat('There are', length(num_df), 'numeric variables, and', length(factor_df), 'factor variables')

for(i in 1:ncol(num_df)){
        if (abs(Skew(num_df[,i]))>0.8){
                num_df[,i] <- log(num_df[,i] +1)
        }
}

Prenum <- preProcess(num_df, method=c("center", "scale"))
print(Prenum)
df_norm <- predict(Prenum, num_df)
dim(df_norm)

DFdummies <- as.data.frame(model.matrix(~.-1, factor_df))
dim(DFdummies)

ZerocolTest <- which(colSums(DFdummies[1:nrow(combined[!is.na(combined$SalePrice),]),])==0)
colnames(DFdummies[ZerocolTest])
DFdummies <- DFdummies[,-ZerocolTest] #removing predictors
fewOnes <- which(colSums(DFdummies[1:nrow(combined[!is.na(combined$SalePrice),]),])<10)
colnames(DFdummies[fewOnes])
DFdummies <- DFdummies[,-fewOnes] #removing predictors
dim(DFdummies)
final <- cbind(num_df, DFdummies)
```



## Transforming SalePrice variable
```{r}
Skew(combined$SalePrice)
qqnorm(combined$SalePrice)
qqline(combined$SalePrice)
layout(mat = matrix(c(1,2),2,1, byrow=TRUE), height = c(1,8))
par(mar=c(0, 3.1, 1.1, 2.1))
boxplot(combined$SalePrice , horizontal=TRUE , xaxt="n", frame=F, main=sprintf('Histogram of SalePrice'))
par(mar=c(4, 3.1, 1.1, 2.1))
hist(combined$SalePrice,main='', xlab = "SalePrice", ylab = "count")
```

```{r}
combined$SalePrice <- log(combined$SalePrice)
Skew(combined$SalePrice)
qqnorm(combined$SalePrice)
qqline(combined$SalePrice)
layout(mat = matrix(c(1,2),2,1, byrow=TRUE), height = c(1,8))
par(mar=c(0, 3.1, 1.1, 2.1))
boxplot(combined$SalePrice , horizontal=TRUE , xaxt="n", frame=F, main=sprintf('Histogram of SalePrice'))
par(mar=c(4, 3.1, 1.1, 2.1))
hist(combined$SalePrice,main='', xlab = "SalePrice", ylab = "count")
```
```{r}
final_df <- final
final_df$SalePrice <- combined$SalePrice
```

## Train and test data
```{r}
library(caret)
set.seed(1)
train_rows <- createDataPartition(y=final_df[,'SalePrice'], list=FALSE, p=.8)
train_dummy <- final_df[train_rows,]
test_dummy <- final_df[-train_rows,]
stopifnot(nrow(train_dummy) + nrow(test_dummy) == nrow(final_df))
write.csv(train_dummy, "../data/processed/train_data_with_dummy.csv")
write.csv(test_dummy, "../data/processed/test_data_with_dummy.csv")

train_rows <- createDataPartition(y=combined[,'SalePrice'], list=FALSE, p=.8)
train <- combined[train_rows,]
test <- combined[-train_rows,]
stopifnot(nrow(train) + nrow(test) == nrow(combined))
write.csv(train, "../data/processed/train_data.csv")
write.csv(test, "../data/processed/test_data.csv")
```

# Modelling

## Ridge regression

```{r}
library(caret)
train_dummy <- read.table("../data/processed/train_data_with_dummy.csv", sep = ",", header = T)
head(train_dummy)
test_dummy <- read.table("../data/processed/test_data_with_dummy.csv", sep = ",", header = T)
head(test_dummy)

custom<- trainControl(method = "repeatedcv",number = 10,repeats = 5,verboseIter = T)
set.seed(1234)
ridge <- train(y=train_dummy$SalePrice, x=train_dummy[,!names(train_dummy) %in% "SalePrice"], method = 'glmnet' , tuneGrid = expand.grid(alpha=0 , lambda =seq(0.001, 0.1, by=0.005))
               ,trControl =custom)
ridge$bestTune
min(ridge$results$RMSE)
a <- predict(ridge,test_dummy[,!names(test_dummy) %in% "SalePrice"])
exp(a)
rmse_rf <- sqrt(mean((exp(test_dummy$SalePrice) -exp(a))^2))
rmse_rf

rss <- sum((exp(a) - exp(test_dummy$SalePrice)) ^ 2)  ## residual sum of squares
tss <- sum((exp(test_dummy$SalePrice) - mean(exp(test_dummy$SalePrice))) ^ 2)  ## total sum of squares
rsq_rf <- 1 - rss/tss
rsq_rf
ridgevar <- varImp(ridge)

```

## Multiple linear regression

### Variable selection

install.packages("randomForest")

http://r-statistics.co/Variable-Selection-and-Importance-With-R.html

## Random forest model
```{r}
library(randomForest)
set.seed(2018)
fit <- randomForest(SalePrice~.,data = train, ntree = 500, mtry = 8, importance = TRUE)
fit
```

### Variable importance plot
```{r}
library(dplyr)
library(ggplot2)
library(randomForest)
varImpPlot(fit)
feat_imp_df<-importance(fit) %>% data.frame() %>% mutate(feature= row.names(.))

ggplot(feat_imp_df, aes(x = reorder(feature,X.IncMSE), 
                         y = X.IncMSE)) +
    geom_bar(stat='identity') +
    coord_flip() +
  labs(y = "% increase in MSE", x = "Features", title = "Feature importance")
```

### Top 20 Important variables
```{r}
imp_vars <-importance(fit)
imp_vars <- sort(imp_vars[,1],decreasing = TRUE)
imp_vars <- imp_vars[1:20]
imp_vars
```

### RF model assesment
```{r}
pred_values = exp(predict(fit,test[,!names(test) %in% "SalePrice"]))
actual_values = exp(test$SalePrice)
rmse_rf <- sqrt(mean((actual_values -pred_values)^2))
rmse_rf

rss <- sum((pred_values - actual_values) ^ 2)  ## residual sum of squares
tss <- sum((actual_values - mean(actual_values)) ^ 2)  ## total sum of squares
rsq_rf <- 1 - rss/tss
rsq_rf
```

<!-- ### Multicollinearity check -->
<!-- ```{r} -->
<!-- library(car) -->
<!-- imp_vars <- imp_vars[!names(imp_vars) %in% "MSSubClass"] -->
<!-- formula <- as.formula(paste("SalePrice", paste(names(imp_vars), collapse = "+"), sep = "~")) -->
<!-- model <- lm(formula, data = combined) -->
<!-- summary(model) -->
<!-- alias(model) -->
<!-- vif(model) -->
<!-- ``` -->

<!-- ### Removing multicollinear variables -->
<!-- ```{r} -->
<!-- imp_vars <- imp_vars[!names(imp_vars) %in% c("X2ndFlrSF","YearRemodAdd" )] -->
<!-- formula <- as.formula(paste("SalePrice", paste(names(imp_vars), collapse = "+"), sep = "~")) -->
<!-- model <- lm(formula, data = combined) -->
<!-- summary(model) -->
<!-- alias(model) -->
<!-- vif(model) -->
<!-- ``` -->

<!-- ### Final Variable subset -->
<!-- ```{r} -->
<!-- final_imp_vars <- names(imp_vars) -->
<!-- final_imp_vars -->
<!-- ``` -->

## Gradient boosting

for gradient boosting tutorial 
https://xgboost.readthedocs.io/en/latest/R-package/xgboostPresentation.html

### Default gradient boosting model
```{r}
label_train <- train_dummy$SalePrice
```

```{r}
library(xgboost)
# train and test data for xgb model
dtrain <- xgb.DMatrix(data = as.matrix(train_dummy[,!names(train_dummy) %in% "SalePrice"]), label = label_train)
dtest <- xgb.DMatrix(data = as.matrix(test_dummy[,!names(test_dummy) %in% "SalePrice"]))
default_param<-list(
        objective = "reg:squarederror",
        booster = "gbtree",
        eta=0.3, #default = 0.3
        gamma=0,
        max_depth=6, #default=6
        min_child_weight=1, #default=1
        subsample=1,
        colsample_bytree=1
)
# cross validation for number of rounds
xgbcv <- xgb.cv( params = default_param, data = dtrain, nrounds = 500, nfold = 5, showsd = T, stratified = T, print_every_n = 40, early_stopping_rounds = 10, maximize = F)
```

we got a train-rmse of 0.014953 and a test-rmse of 0.143685.

## Tuning hyperparameters
```{r}
set.seed(1)
grid = expand.grid(
nrounds = 50,
eta = c(0.1, 0.05, 0.01),
max_depth = c(2, 3, 4, 5, 6),
gamma = 0,
colsample_bytree=1,
min_child_weight=c(1, 2, 3, 4 ,5),
subsample=1
)
cv_control <-trainControl(method="cv", number=5)
calibration_model <- train(x=train_dummy[,!names(train_dummy) %in% "SalePrice"], y=train_dummy$SalePrice, method='xgbTree', trControl= cv_control, tuneGrid=grid) 

```

```{r}
calibration_model$bestTune
```

```{r}
default_param<-list(
        objective = "reg:squarederror",
        booster = "gbtree",
        eta=0.1, #default = 0.3
        gamma=0,
        max_depth=5, #default=6
        min_child_weight=2, #default=1
        subsample=1,
        colsample_bytree=1
)
# cross validation for number of rounds
xgbcv <- xgb.cv( params = default_param, data = dtrain, nrounds = 500, nfold = 5, showsd = T, stratified = T, print_every_n = 40, early_stopping_rounds = 10, maximize = F)
```

After tuning the model parameters we got train-rmse: 0.047392 and test-rmse: 0.137014 which is an improvement on the test-rmse when compared to the default training prarmeter values.

```{r}
xgb_model <- xgb.train(data = dtrain, params=default_param, nrounds = 61)
XGBpred <- exp(predict(xgb_model, dtest))
```

```{r}
actual <- exp(test_dummy$SalePrice)
rmse <- sqrt(mean((actual -XGBpred)^2))
rmse

rss <- sum((XGBpred - actual) ^ 2)  ## residual sum of squares
tss <- sum((actual - mean(actual)) ^ 2)  ## total sum of squares
rsq <- 1 - rss/tss
rsq
head(data.frame(actual,XGBpred))
```

# Results

